---
title: "Predicting Breast cancer"
author: "Titus Njonjo"
date: "5/22/2020"
output: html_document
---

## This is **an analysis of breast cancer** data using decision tree algorithm to predict if a patient has breast cancer or not. the data has the following variables:
* class  * age  * menopause  * tumour size  * inv nodes	* nodes cap	 * deg-malig * breast  * breast-quad
* irradiat

##Loading required packages 
```{r  message= FALSE, Warning=FALSE, loading required packages }
library(rpart)
library(rpart.plot)
library(readr)
library(rattle)
library(caret)
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)
```
## the dataset can be described as majorly as categorical data
```{r message=FALSE,warning=FALSE ,importing the dataset}
breast <- read_csv("C:/Users/LAIKAM-TITO/desktop/MY PROJECT/breast-cancer.csv")
breast <- mutate_if(breast,is.character,as.factor)
str(breast)
```
## Checking if the data is complete.From the results , it is with no missing values 
```{r}
table(complete.cases(breast))
```
```{r}
dataentry(breast,)

```

## Splitting the data set for training and testing
```{r}
set.seed(1000)
split <- sample(seq_len(nrow(breast)), size=floor(0.8*nrow(breast)))
train <- breast[split,]
testing <- breast[-split,]
train
testing
```

## At this stage , i will be doing an exploratory analysis from the data, where I wish to answer the following questions from the data:
* which age-group has high likelihood to be diagonised with breast cancer?
* Which breast is likely to be affected by cancerous cells ?
* which section of the breast is likely to be affected by the cancerous cells ?

## Q1.which age-group has high likelihood to be diagonised with breast cancer?
```{r Exploratory data analysis}
qplot(class, data=train, fill=age)+facet_grid(.~age)+theme(axis.text.x = element_text(angle=90, vjust=0.9))+labs(title="Women in the 50-59 age bracket are likely to be diagonised with breast cancer,
while women in the 20-29 and 70-79 age bracket 
are less likely to be diagonised",
  y="Count",
 caption = "Source:Institute of Oncology")
ggsave("agegroup.pdf",)
```
## Q2. Which breast is likely to be affected by cancerous cells ?
```{r}
qplot(class, data=train, fill=breast)+facet_grid(.~breast)+labs(title = "The left breast is more likely to be affected by cancerous cell ",
y="Count",
caption = "Source: Institute of Oncology ")

```
## Q3.Which section of the breast is likely to be affected by the cancerous cells ?
```{r}
qplot(class, data=train, fill= `breast-quad`)+facet_grid(.~`breast-quad`)+theme(axis.text.x=element_text(angle=90,vjust=0.9))
```
##Q4. Which group of women are more likely tobe diagonised with breast cancer?
```{r}
qplot(class, data=breast, fill=menopause)+facet_grid(.~menopause)+theme(axis.text.x=element_text(angle=90,vjust=0.9))+ labs(title = "Women in the premenopause group are  likely
to be diagonised with breast cancer", y="count")
```
##Q5. Does tumor size indicate cancer?
```{r}
qplot(class, data=breast, fill=`tumor-size`)+facet_grid(.~`tumor-size`)+theme(axis.text.x=element_text(angle=90,vjust=0.9))+ labs(title = "Women in the premenopause group are  likely
to be diagonised with breast cancer", y="count")
```
```{r SPLITTING FEATURES}
number.perfect.splits <- apply(X=breast[-1], MARGIN = 2, FUN = function(col){
t <- table(breast$class,col)
sum(t == 0)
})
order <- order(number.perfect.splits,decreasing = TRUE)
number.perfect.splits<- number.perfect.splits[order]
par(mar=c(10,4,3,3))
barplot(number.perfect.splits,
main="Number of perfect splits vs feature",
xlab="",ylab="Feature",las=2,col="wheat")
```

```{r modeling}
penalty.matrix <- matrix(c(0,1,10,0), byrow=TRUE, nrow=2)
penalty.matrix
```
```{r}
tree <- rpart(class~. , data=train,parms=list(loss=penalty.matrix), method="class")
tree
```
##Ploting a decision tree
```{r}
rpart.plot(tree,nn=TRUE)
```
##Making prediction out of the model
```{r}
pred1 <- predict(object=tree,testing[-1],type="class")
write_file(pred1,"C:/Users/LAIKAM-TITO/Desktop",)
```
```{r}

```
```{r}
h <- table(testing$class,pred1) 
confusionMatrix(h)
h
```

